---
title: "In Class Assignment-1"
editor: visual
---

Using nycflights13 package I have prepared the following analyses.

```{r}
library(tidyverse)
library(nycflights13)
```

## Analysis-1 Number of tailnumbers per grouped number of seats.

```{r}
planes %>% 
  mutate(seat_bracket=case_when(seats<=100 ~ "(0-100]", seats<=200 ~"(100-200]",seats<=300 ~"(200-300]", seats<=400 ~"(300-400]",TRUE ~ "(400-HI]")) %>% 
  group_by(seat_bracket) %>%
  summarise(count=n()) %>% 
  print(n=Inf)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

## Analysis-2 Average, min and max speed with respect to type, engine and number of engines.

First, we check the distinct values of speed column.

```{r}
planes %>%
  group_by(speed) %>%
  summarise(count=n()) %>%
  arrange(desc(speed)) %>%
  print(n=Inf)
```

We see that most of the speed data is non-existent. With the remaining data here is the table for average, min and max of speed by type, engine and number of engines.

```{r}
options(dplyr.summarise.inform=FALSE)
planes %>%
  filter(!is.na(speed)) %>%
  group_by(type,engine,engines) %>%
  summarise(avg_speed=round(mean(speed),2),min_speed=min(speed),max_speed=max(speed)) %>%
  arrange(type,engine,engines) %>%
  print(n=Inf)
```
